Analytic Occultation Light Curves in Reflected Light

### 2020-04-14 tex/ms.tex
finished qT

if $F$ is the antiderivative of some function $f$, we may use the

@@

     & = \Delta \mathbf{F}(\pmb{\varphi})
where $\mathbf{F}$ is the vector given by
    \mathbf{F}(\pmb{\varphi}) =
    \bigg( F(\varphi_0), F(\varphi_1), F(\varphi_2), F(\varphi_3),
    {\cdot\cdot\cdot}, F(\varphi_{N - 2}), F(\varphi_{N - 1}) \bigg)^\top

@@

\subsection{The integral along the occultor limb, $\pT$}
\label{sec:pT}
%
In this section we present a solution to Equation~(\ref{eq:pT}). The first
order of business is to derive expressions for the integration limits
$\pmb{\phi}$. Depending on the integration case, these limits will correspond
to the point of intersection between the limb of the occultor and the
limb of the occulted body and/or the point of intersection between the limb
of the occultor and the terminator of the occulted body.
The former is given by
\citep[c.f. Equation~24 in][]{Luger2019}
%
\begin{align}
    \phi_0 & = \frac{\pi}{2} \pm \left(\arcsin\left(\frac{1 - r_o ^ 2 - b_o ^ 2}{2 b_o r_o}\right) - \frac{\pi}{2}\right)
    \quad,
\end{align}
%
where the sign is chosen such that the point
$(r_o\cos\phi_l, b_o + r_o\sin\phi_l)$ is on the dayside of the occulted body,
and the latter (of which there may be multiple) is given by
%
% Introduce the function
% %
% \begin{proof}{}
%     \textrm{\faAdjust}(x, y) &=
%     \begin{cases}
%         \textrm{\faCircleO} & \qquad -x \sin\theta' + y\cos\theta' \ge b \sqrt{1 - (x\cos\theta' + y\sin\theta') ^ 2}
%         \\
%         \textrm{\faCircle}  & \qquad \text{otherwise}
%     \end{cases}
% \end{proof}
% %
% which returns \faCircleO\ if point $(x, y)$ is on the dayside of the body and
% \faCircle\ if it is on the nightside.
%
%
\begin{align}
    \pmb{\phi_1} & =
    \theta' +
    \atantwo
    \left(b\sqrt{1 - {\mathbf{x''}}^2 - y''_o}, \mathbf{x''} - x''_o\right)
    \quad
\end{align}
%
where $\mathbf{x''}$ are the roots of the quartic (Equation~\ref{eq:quartic}).
These angles are then sorted and placed in the vector $\pmb{\phi}$ such that
the integration is always performed in a counter-clockwise sense about
the center of the integration region.
%
The left panel in Figure~\ref{fig:geometry} shows a configuration in which
the lower integration limit $\phi_0 = 158.2^\circ$ corresponds to the
point of intersection between the limbs of the two bodies and the upper
integration limit $\phi_1 = 223.3^\circ$ corresponds to the
limb-terminator intersection. Both angles are measured counter-clockwise
from the $x'$-axis.

@@

\subsection{The integral along the terminator, $\tT$}
%
In this section we present a solution to Equation~(\ref{eq:tT}), the
line integral along the day/night terminator of the occulted body. As
before, the first thing we must do is derive expressions for the integration
limits $\pmb{\xi}$.
%
Depending on the integration case, these limits
will corresponds to the point of intersection between the terminator and
the limb of the occultor and/or the point of intersection between the
terminator and the limb of the occulted body. The former is given by
%
\begin{proof}{}
    \pmb{\xi_0} &= \atantwo\left(\sqrt{1 - {\mathbf{x''}}^2}, \mathbf{x''} \right)
\end{proof}
%
where $\mathbf{x''}$ are the roots of the quartic (Equation~\ref{eq:quartic}).
The latter is given by
%
\begin{proof}{}
    \xi_1 &=
    \begin{cases}
        0   & \qquad \qquad (1 - x_o'')^2 + {y_o''}^2 < r_o^2
        \\
        \pi & \qquad \qquad \text{otherwise}
        \quad.
    \end{cases}
\end{proof}
%
As before, these angles are then sorted and placed in the vector
$\pmb{\xi}$ such that the integration is performed counter-clockwise
about the center of the integration region.
%
The middle panel of Figure~\ref{fig:geometry} shows a case where
$\xi_0 = 96.5^\circ$ corresponds to the point of intersection between the occultor
limb and the terminator and $\xi_1 = 0^\circ$ corresponds to the
point where the terminator extends onto the backside of the body.
%
Note, importantly, that unlike $\pmb{\phi}$, the angle $\pmb{\xi}$ is not
measured between the $x$-axis and a point on the curve of $\mathcal{T}$.
Recall that $\pmb{\xi}$ is an angular parameter of the ellipse, so it is
measured in the same way as the eccentric anomaly in a Keplerian orbit:
it is the
angle between the $x$-axis and the perpendicular
projection of a point on the ellipse onto the unit circle
(see Figure~\ref{fig:geometry}).

@@

        {_1}Z^{\mu,\nu}_{j,k}(b, \theta')

@@

         & \qquad \mu, \nu, \, \text{even}

@@

        \Bigg(

@@

        \Bigg)

@@

        {_3}Z^{\mu,\nu}_{j}(b, \theta')

@@

         & \qquad \nu \, \text{odd}, \mu = 1, l \, \text{even}

@@

        \sin\theta'{_4}Z^{\mu,\nu}_{j}(b, \theta')

@@

        b\cos\theta'{_4}Z^{\mu,\nu}_{j}(b, \theta')

@@

@@

         & \qquad \nu \, \text{odd}, \mu = 1, l \, \text{odd}

@@

        {_2}Z^{\mu,\nu}_{j,k}(b, \theta')

@@

         & \qquad \text{otherwise}

@@

    {_1}Z^{\mu,\nu}_{j,k}(b, \theta') & =
    \binom{\frac{\nu}{2}}{j}
    \binom{\frac{\mu + 2}{2}}{k}
    (-1)^{\frac{\mu + 2 - 2k}{2}}
    b^{l + 1 - j - k}
    \sin^\frac{\mu + 2 + 2j - 2k}{2}\theta'
    \cos^\frac{\nu - 2j + 2k}{2}\theta'
    {_2}Z^{\mu,\nu}_{j,k}(b, \theta') & =
    \binom{\frac{\nu - 1}{2}}{j}
    \binom{\frac{\mu - 3}{2}}{k}
    (-1)^{\frac{\mu - 3 - 2k}{2}}
    b^{l - 2 - j - k}
    b_c^3
    \sin^\frac{\mu - 3 + 2j - 2k}{2}\theta'
    \cos^\frac{\nu - 1 - 2j + 2k}{2}\theta'
    {_3}Z^{\mu,\nu}_{j}(b, \theta')   & =
    \binom{l - 2}{j}
    (-1)^{l - 2 - j}
    b^{l - 2 - j}
    b_c^3
    \sin^{l - 2 - j}\theta'
    \cos^j\theta'
    {_4}Z^{\mu,\nu}_{j}(b, \theta')   & =
    \binom{l - 3}{j}
    (-1)^{l - 3 - j}
    b^{l - 3 - j}
    b_c^3
    \sin^{l - 3 - j}\theta'
    \cos^j\theta'

@@

@@

\appendix

\xxx{Migrate this up.}

%
The function $\mathcal{H}_{u,v}$ is given by
%
\begin{align}
    \label{eq:H}
    \mathcal{H}_{u,v}(\vlambda) & =
    \vint{\vlambda}{
        \cos^u\varphi
        \sin^v\varphi
    }
    \quad.
\end{align}
%
The integral in this expression is the same as that in Equation (D27)
of \citet{Luger2019}, except for a change in the limits of integration.
We can compute this integral recursively given four lower boundary conditions:
%
\begin{proof}{H}
    \label{eq:Hlower}
    \mathcal{H}_{0,0}(\vlambda) &= \Delta \vlambda
    %
    \nonumber \\
    %
    \mathcal{H}_{1,0}(\vlambda) &= \Delta \sin\vlambda
    %
    \nonumber \\
    %
    \mathcal{H}_{0,1}(\vlambda) &= -\Delta \cos\vlambda
    %
    \nonumber \\
    %
    \mathcal{H}_{1,1}(\vlambda) &= -\frac{\Delta\cos^2\vlambda}{2}
    %
    \quad.
\end{proof}
%
The remaining terms may be computed by upward recursion using the
relations
%
\begin{proof}{H}
    \label{eq:Hrec1}
    \mathcal{H}_{u,v}(\vlambda) &=
    \frac{
        -\Delta \left(
        \cos^{u + 1} \vlambda
        \sin^{v - 1} \vlambda
        \right)
        +(v - 1)\mathcal{H}_{u,v - 2}(\vlambda)
    }{u + v}
\end{proof}
%
for $u < 2, v \ge 2$ and
%
\begin{proof}{H}
    \label{eq:Hrec2}
    \mathcal{H}_{u,v}(\vlambda) &=
    \frac{
        \Delta \left(
        \cos^{u - 1} \vlambda
        \sin^{v + 1} \vlambda
        \right)
        + (u - 1)\mathcal{H}_{u - 2,v}(\vlambda)
    }{u + v}
\end{proof}
%
for all remaining terms.

\subsection{$\mathcal{\mathcal{T}}_{2}$}
%
The function $\mathcal{\mathcal{T}}_{2}$ is given by
%
\begin{align}
    \label{eq:T2}
    \mathcal{\mathcal{T}}_{2}(b, \vxi) & = ?
    \quad.
\end{align}

\begin{proof}{}
    \mathcal{\mathcal{T}}_{2}(b, \vxi) &= -\sgn(b) \Delta \mathbf{F}(b, \vxi)
\end{proof}
%
where
%
\begin{proof}{}
    \begin{split}
        f_i(b, \xi_i) =
        \frac{1}{3}\scalebox{1.4}{\Bigg(}
        &\arctan\left( \frac{|b|\sin\xi_i}{\cos\xi_i} \right)
        %
        \\
        %
        &- \sgn\left({\sin\xi_i}\right)
        \left(
        \arctan
        \left(
            \frac{
                \left(\frac{\sin\xi_i}{1 + \cos\xi_i}\right)^2 + 2 b^2 - 1
            }{2 |b| \sqrt{1 - b^2}}
            \right)
        + |b| \sqrt{1 - b^2} \cos\xi_i
        \right)
        %
        \\
        %
        &+ \phi(b, \xi_i)
        \scalebox{1.4}{\Bigg)}
    \end{split}
\end{proof}
%
and
%
\begin{proof}{}
    \phi(b, \xi_i) & =
    \begin{cases}
        0                          & \qquad 0 \leq \xi_i < \frac{\pi}{2}     \\
        \pi                        & \qquad \frac{\pi}{2} \leq \xi_i < \pi   \\
        2 |b| \sqrt{1 - b^2}       & \qquad \pi < \xi_i \leq \frac{3\pi}{2}  \\
        \pi + 2 |b| \sqrt{1 - b^2} & \qquad \frac{3\pi}{2} \leq \xi_i < 2\pi
        \quad.
    \end{cases}
\end{proof}

>>>>

if $g$ is the antiderivative of some function $f$, we may use the

@@

     & = \Delta \mathbf{g}(\pmb{\varphi})
where $\mathbf{g}$ is the vector given by
    \mathbf{g}(\pmb{\varphi}) =
    \bigg( g(\varphi_0), g(\varphi_1), g(\varphi_2), g(\varphi_3),
    {\cdot\cdot\cdot}, g(\varphi_{N - 2}), g(\varphi_{N - 1}) \bigg)^\top

@@

\subsection{The integral along the occultor limb, $\pT$}
\label{sec:pT}
%
In this section we present a solution to Equation~(\ref{eq:pT}). The first
order of business is to derive expressions for the integration limits
$\pmb{\phi}$. Depending on the integration case, these limits will correspond
to the point of intersection between the limb of the occultor and the
limb of the occulted body and/or the point of intersection between the limb
of the occultor and the terminator of the occulted body.
The former is given by
\citep[c.f. Equation~24 in][]{Luger2019}
%
\begin{align}
    \phi_0 & =
    \frac{\pi}{2} \pm \left(\arcsin\left(\frac{1 - r_o ^ 2 - b_o ^ 2}{2 b_o r_o}\right) - \frac{\pi}{2}\right)
    \quad,
\end{align}
%
where the sign is chosen such that the point
$(r_o\cos\phi_0, b_o + r_o\sin\phi_0)$ is on the dayside of the occulted body,
and the latter (of which there may be multiple) is given by
%
% Introduce the function
% %
% \begin{proof}{}
%     \textrm{\faAdjust}(x, y) &=
%     \begin{cases}
%         \textrm{\faCircleO} & \qquad -x \sin\theta' + y\cos\theta' \ge b \sqrt{1 - (x\cos\theta' + y\sin\theta') ^ 2}
%         \\
%         \textrm{\faCircle}  & \qquad \text{otherwise}
%     \end{cases}
% \end{proof}
% %
% which returns \faCircleO\ if point $(x, y)$ is on the dayside of the body and
% \faCircle\ if it is on the nightside.
%
%
\begin{align}
    \pmb{\phi_1} & =
    \theta' +
    \atantwo
    \left(b\sqrt{1 - {\mathbf{x''}}^2 - y''_o}, \mathbf{x''} - x''_o\right)
    \quad
\end{align}
%
where $\mathbf{x''}$ are the roots of the quartic (Equation~\ref{eq:quartic}).
These angles are then wrapped to the range $[0, 2\pi)$
and sorted into the vector $\pmb{\phi}$ such that
the integration is always performed in a counter-clockwise sense about
the center of the integration region.
%
The left panel in Figure~\ref{fig:geometry} shows a configuration in which
the lower integration limit $\phi_0 = 158.2^\circ$ corresponds to the
point of intersection between the limbs of the two bodies and the upper
integration limit $\phi_1 = 223.3^\circ$ corresponds to the
limb-terminator intersection. Both angles are measured counter-clockwise
from the line $x' = b_o$.

@@

\subsection{The integral along the terminator, $\tT$}
%
In this section we present a solution to Equation~(\ref{eq:tT}), the
line integral along the day/night terminator of the occulted body. As
before, the first thing we must do is derive expressions for the integration
limits $\pmb{\xi}$.
%
Depending on the integration case, these limits
will corresponds to the point of intersection between the terminator and
the limb of the occultor and/or the point of intersection between the
terminator and the limb of the occulted body. The former is given by
%
\begin{proof}{}
    \pmb{\xi_0} &=
    \atantwo\left(\sqrt{1 - {\mathbf{x''}}^2}, \mathbf{x''} \right)
\end{proof}
%
where $\mathbf{x''}$ are the roots of the quartic (Equation~\ref{eq:quartic}).
The latter is given by
%
\begin{proof}{}
    \xi_1 &=
    \begin{cases}
        0   & \qquad \qquad (1 - x_o'')^2 + {y_o''}^2 < r_o^2
        \\
        \pi & \qquad \qquad \text{otherwise}
        \quad.
    \end{cases}
\end{proof}
%
As before, these angles are then
wrapped to the range $[0, 2\pi)$ and
sorted into the vector
$\pmb{\xi}$ such that the integration is performed counter-clockwise
about the center of the integration region.
%
The middle panel of Figure~\ref{fig:geometry} shows a case where
$\xi_0 = 96.5^\circ$ corresponds to the point of intersection between the occultor
limb and the terminator and $\xi_1 = 0^\circ$ corresponds to the
point where the terminator extends onto the backside of the body.
%
Note, importantly, that unlike $\pmb{\phi}$, the angle $\pmb{\xi}$ is not
measured between the horizontal and a point on the curve of $\mathcal{T}$.
Recall that $\pmb{\xi}$ is an angular parameter of the ellipse, so it is
measured in the same way as the eccentric anomaly in a Keplerian orbit:
it is the
angle between the semi-major axis of the ellipse and the perpendicular
projection of a point on the ellipse onto the unit circle
(see Figure~\ref{fig:geometry}).

The solution to Equation~(\ref{eq:tT}) involves repeated application
of the binomial theorem. As before, we simply present the result, with
validation links:

@@

        Z^{\frac{\nu}{2}, \frac{\mu+2}{2}}_{j,k}
        (b, \theta') \,

@@

        \mathrel{\raisebox{1.25em}{$\mu, \nu, \, \text{even}$}}

@@

        \Big\{

@@

        \Big\}

@@

        Z^{0, l-2}_{0,k}
        (b, \theta') \,

@@

        \mathrel{\raisebox{1.25em}{$\nu \, \text{odd}, \mu = 1, l \, \text{even}$}}

@@

        \sin\theta'
        Z^{0, l-3}_{0,k}
        (b, \theta') \,

@@

        \, b\cos\theta'
        Z^{0, l-3}_{0,k}
        (b, \theta') \,

@@

        \mathrel{\raisebox{1.25em}{$\nu \, \text{odd}, \mu = 1, l \, \text{odd}$}}

@@

@@

        Z^{\frac{\nu-1}{2}, \frac{\mu - 3}{2}}_{j,k}
        (b, \theta') \,

@@

        \mathrel{\raisebox{1.25em}{$\text{otherwise}$}}

@@

where we define
    Z^{u,v}_{j,k}(b, \theta') & =
    \binom{v}{k}
    (-1)^{v-k}
    b^{u+v-j-k}
    \sin^{v+j-k}\theta'
    \cos^{u-j+k}\theta'

@@

The components of the vector $\viH$ are given by the integral
    \label{eq:H}
    \iH_{u,v}(\vxi) & =
    \vint{\vxi}{
        \cos^u\varphi
        \sin^v\varphi
This integral is the same as that in Equation (D27)
of \citet{Luger2019}, except for a change in the limits of integration.
We can compute this integral recursively given four lower boundary conditions:
\begin{proof}{H}
    \label{eq:Hlower}
    \iH_{0,0}(\vxi) &= \Delta \vxi
    \iH_{1,0}(\vxi) &= \Delta \sin\vxi
    \iH_{0,1}(\vxi) &= -\Delta \cos\vxi
    \iH_{1,1}(\vxi) &= -\frac{\Delta\cos^2\vxi}{2}
The remaining terms may be computed by upward recursion using the
relations
\begin{proof}{H}
    \label{eq:Hrec1}
    \iH_{u,v}(\vxi) &=
        -\Delta \left(
        \cos^{u + 1} \vxi
        \sin^{v - 1} \vxi
        +(v - 1)\iH_{u,v - 2}(\vxi)
    }{u + v}
for $u < 2, v \ge 2$ and
\begin{proof}{H}
    \label{eq:Hrec2}
    \iH_{u,v}(\vxi) &=
        \Delta \left(
        \cos^{u - 1} \vxi
        \sin^{v + 1} \vxi
        + (u - 1)\iH_{u - 2,v}(\vxi)
    }{u + v}
for all remaining terms.

\subsection{The integral along the occulted body limb, $\qT$}
%
The final line integral we must solve is the integral along the boundary
of the occulted body,  Equation~(\ref{eq:qT}). Fortunately, this is
also the easiest of the three.
%
The limits of integration $\vlambda$
correspond to the point at which the terminator crosses from the dayside
to the night side,
%
\begin{proof}{}
    \lambda_0 &=
    \begin{cases}
        \theta'        & \qquad \qquad \cos^2\theta' + (\sin\theta' - b_o)^2 < r_o^2
        \\
        \theta ' + \pi & \qquad \qquad \text{otherwise}
        \quad.
    \end{cases}
\end{proof}
%
and the point of intersection between the limb of the occultor
and the limb of the occulted body,
%
\begin{proof}{}
    \lambda_1 & =
    \frac{\pi}{2} \pm \left(\arcsin\left(\frac{1 - r_o ^ 2 + b_o ^ 2}{2 b_o}\right) - \frac{\pi}{2}\right)
    \quad,
\end{proof}
%
where the sign is chosen such that the point
$(\cos\lambda_1, \sin\lambda_1)$ is on the dayside of the occulted body.
%
As before, these angles are
wrapped to the range $[0, 2\pi)$ and
placed in the vector
$\lambda$ such that the line integral is taken in the counter-clockwise
direction about the center of the integration region. The right panel
of Figure~\ref{fig:geometry} shows a case where
$\lambda_0 = 75^\circ$ and $\lambda_1 = 130.5^\circ$. Both angles are
measured counter-clockwise from the $x'$-axis.

Given $\vlambda$, the solution to Equation~(\ref{eq:qT}) is straightforward:
%
\begin{proof}{}
    \mathbb{q}_n &=
    \begin{cases}
        \iH_{\frac{\mu + 4}{2}, \frac{\nu}{2}}(\vlambda)
                                  & \qquad \mu, \nu, \, \text{even}
        \\[1em]
        \frac{1}{3}\Delta\vlambda & \qquad \mu = \nu = 1
        \\[1em]
        0                         & \qquad \text{otherwise}
        \quad,
    \end{cases}
\end{proof}
%
where the vector $\viH$ is given by Equation~(\ref{eq:H}).

@@
